Modularizing Street Map Generation \\ \vspace{8pt} \large An Adventure in Tensors

### 2013-12-12 paper/paper.tex
references and figs in underlying math

\documentclass[twocolumn]{article}
\usepackage{amsmath}

@@

\section{Mathematical Background}
A tensor $t$ is a *magical thing* represented by a 2x2 symmetric, traceless
matrix:
\[
    \sqmat{\cos{2\theta}}{\sin{2\theta}}{\sin{2\theta}}{-\cos{2\theta}}
    = \sqmat{a}{b}{b}{-a}
\]
where $R$ is the magnitude and $\theta$ is the direction, and
$R\geq0 \textrm{ and } \theta\in[0,2\pi)$. Since the matrix is symmetric and
traceless, it is only necessary to store two values, $a$ and $b$ to encode
the tensor.

@@

@@

>>>>

\documentclass[twocolumn]{article}
\usepackage{amsmath}
\usepackage{graphicx}

%borders
\usepackage{float}
\floatstyle{boxed}
\restylefloat{figure}

@@

\section{Mathematical Background}
\label{sec:math}
A tensor $t$ is a *magical thing* represented by a 2x2 symmetric, traceless
matrix:
\[
    \sqmat{\cos{2\theta}}{\sin{2\theta}}{\sin{2\theta}}{-\cos{2\theta}}
    = \sqmat{a}{b}{b}{-a}
\]
where $R$ is the magnitude and $\theta$ is the direction, and
$R\geq0 \textrm{ and } \theta\in[0,2\pi)$. Since the matrix is symmetric and
traceless, it is only necessary to store two values, $a$ and $b$ to encode
the tensor.

@@

\begin{figure*}[t!]
\begin{verbatim}
-- returns a tuple of the (larger, smaller) eigenvectors of a tensor
eigenvalues :: Tensor -> (Float, Float)
eigenvalues (Tensor a b d) =
  let eval = sqrt (a*a + b*b)
  in  (eval, (-1)*eval)

-- returns a tuple of the (major, minor) eigenvectors for a tensor
eigenvectors :: Tensor -> (V2.Vector2, V2.Vector2)
eigenvectors (Tensor a 0.0 d) =
  if a > d then (V2.Vector2 1 0, V2.Vector2 0 1)
     else (V2.Vector2 0 1, V2.Vector2 1 0)
     eigenvectors (Tensor a b d) =
       let (eval1, eval2) = eigenvalues (Tensor a b d)
           mkEvec eval    = V2.Vector2 1 ((eval - a) / b)
       in (mkEvec eval1, mkEvec eval2)
\end{verbatim}
\caption{Haskell implementation of eigenvector calculation}
\label{fig:evecs}
\end{figure*}

@@

An example of this can be seen in the calculation of eigenvectors. We derived
the formula for the eigenvector of a symmetric, traceless tensor 
(see Section~\ref{sec:math}) and implemented it in the functions in
Figure~\ref{fig:evecs}.

In [Chen], there is a 75-line function,
\texttt{cal\_eigenvecs\_onevert\_quad}, to do approximately the same thing. There
is a fair amount of documentation in the body of the function, explaining what
each line does, but there also seem to be some side effects relating to global
variables and large data structures that the authors found difficult to follow,
mostly by the nature of the language.  Generally speaking, we found that the
mathematics was significantly easier to express and review in a Haskell
implementation.
